---
title: "Initial analysis"
author: César Soto Valero
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    css: main.css
    number_sections: true
    toc: false
    toc_depth: 3
    toc_float: 
      collapsed: false
---


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Import the header 
source("header.R")
# Clear the R environment
rm(list=ls())
```


```{r echo=FALSE}
# Read the data
raw_results <- read.delim("Data/results.csv", sep = ",", header = T, stringsAsFactors = TRUE)
```


# Legend

- **AUC**: Always Used Classes
- **ABC**: Always Bloated Classes
- **AUM**: Always Used Methods
- **ABM**: Always Used Methods
- **UTABC**: Used Then Always Bloated Classes
- **UTABM**: Used Then Always Bloated Methods
- **BTAUC**: Bloated Then Always Used Classes
- **BTAUM**: Bloated Then Always Used Methods
- **OCSC**: Oscillated Classes
- **OCSM**: Oscillated Methods

## Bloat evolution results

```{r echo=FALSE, warning=FALSE}
result_plot <- raw_results %>%
  ggplot(aes(x = Lib, y = Count, fill = Metric, label = Count)) +
  geom_bar(position="stack", stat="identity", width = 0.75) +
  # geom_text(aes(label=Count), size = 2, position=position_dodge(width=0.75), hjust=-0.25) +
  scale_fill_manual(values = c( "#00cd6c", "#ffc61e", "#a0b1ba", "#ff1f5d", "#00FFFF", "#af58ba", "#f28522", "#a6761d", "#000000", "#87ceeb")) +
  coord_flip() +
    theme(
      legend.position="right",
      # axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        # panel.grid.minor.x = element_blank()
        ) 

# green: #00cd6c
# yellow: #ffc61e
# gray: #a0b1ba
# red: #ff1f5d
# blue: #009ade
# purple: #af58ba
# orange: #f28522
# brown: #a6761d

# PLOT
result_plot
# save_figure(result_plot, "Figures/result_plot", 8, 2)
```


## Aggregated results per metric

```{r echo=FALSE, warning=FALSE}
result_plot <- raw_results %>%
  ggplot(aes(x = Metric, y = Count, fill = Metric, label = Count)) +
  geom_bar(position="stack", stat="identity", width = 0.75) +
  # geom_text(aes(label=Count), size = 2, position=position_dodge(width=0.75), hjust=-0.25) +
  scale_fill_manual(values = c( "#00cd6c", "#ffc61e", "#a0b1ba", "#ff1f5d", "#00FFFF", "#af58ba", "#f28522", "#a6761d", "#000000", "#87ceeb")) +
  coord_flip() +
    theme(
      legend.position="right",
      # axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        # panel.grid.minor.x = element_blank()
        ) 

result_plot
```

## Number of versions considered per library

```{r echo=FALSE, warning=FALSE}
result_plot <- raw_results %>%
  ggplot(aes(x = Lib, y = NbVersions, fill = Lib)) +
  geom_bar(position="dodge", stat="identity", width = 0.75) +
  scale_fill_manual(values = c( "#00cd6c", "#ffc61e", "#a0b1ba", "#ff1f5d", "#00FFFF", "#af58ba", "#f28522", "#a6761d", "#000000", "#87ceeb")) +
  coord_flip() +
    theme(
       legend.position="right",
      # axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        # panel.grid.minor.x = element_blank()
        ) 

result_plot
```
